Network based multifactorial modelling microRNA-target interaction.

### 2019-08-22 article.tex
article files

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Network based multifactorial modelling microRNA-target interaction.},
            pdfauthor={Selcen Ari; Alper Yilmaz},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Network based multifactorial modelling microRNA-target interaction.}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Selcen Ari\footnote{Department of Bioengineering, Yildiz Technical
  University, Istanbul, Turkey} \\ Alper Yilmaz\footnote{Department of Bioengineering, Yildiz Technical
  University, Istanbul, Turkey}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{09 Ağustos 2019}

@@

Competing endogenous RNA (ceRNA) regulations and crosstalk between
various types of non-coding RNA in human is remarkable in means of miRNA
regulation. Many studies have pointed out that an alteration in
miRNA:target interaction can result in unexpected changes due to
indirect and complex interactions. In this paper, we defined a new
network-based model that handles miRNA:ceRNA interactions with
expression values. Our model is able to handle miRNA interaction factors
such as seed type, binding energy, if provided. Our approach is able to
reveal that a perturbation in an element of network affects whole
competing elements differently and cooperative efficiencies of miRNAs on
common targets could be calculated. Our findings emphasized importance
of miRNA:target ratios being crucial, as reported by previous studies.
We have showed that the competing elements which have the same or close
expression values may not be affected equally from the perturbation
because of repression functionality depended on interaction factors of
miRNA target pairs (TOFIX: Hocam etkileşim faktörlerinden dolayı
ekspresyonları yakın olsa bile baskılanma aktitesinin aynı
sergilenmeyeceğini ifade etmeye çalıştım. Bu şekilde anlaşılıyor mu?).
We applied the model to real sample consisting of breast cancer gene and
miRNA expression dataset and experimental miRNA:target interaction
dataset all generated via high throughput sequencing methods. A gene
over-expressed in tumor tissue, namely \emph{ABCC1}, is used as
perturbing element. We have observed that change in expression level of
single gene in miRNA:target network is sufficient to perturb regulations
in whole network, due to unforeseen and unpredicted regulation which are
only visible when considered in network context. Therefore, this model
helps unveiling the crosstalk between elements in miRNA:target network
where abundance of target and sponge effect are taken into account. The
model is scalable and can be plugged in with emerging miRNA effectors
such as circRNAs. The model is available as R package
\href{https://github.com/selcenari/ceRNAnetsim}{ceRNAnetsim}.

@@

MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key
regulator of gene expression through various post-transcriptional
mechanisms. Although the mechanisms by which miRNA represses are not
fully understood, miRNAs predominantly repress their targets. Repressive
activities of miRNAs vary depending on many factors that are significant
to microRNA:target interactions. These factors include microRNA:target
binding energy, binding location in target sequence, base pairing types
between miRNA and target, abundance of miRNAs and targets (Grimson et
al. 2007). Binding energies of miRNA:target complexes vary based on
nucleotide context and determine folding stability of complex (Cao and
Chen 2012). It has been demonstrated that the binding energy between
miRNA and target indicates stability or affinity of complex (Helwak et
al. 2013) and does not directly determine repressive activity of miRNA
(Cao and Chen 2012). Early studies have argued that 2-8 nt sequence,
seed, located in miRNA 5'end bind to specific sequence located in 3'UTR
of its target (Bartel 2004; Lewis, Burge, and Bartel 2005). In recent
studies, it has been shown that miRNAs can interact with targets via
sequences located in regions such as 5'UTR or CDS (Hausser et al. 2013;
Helwak et al. 2013; Moore et al. 2015). These studies also showed that
binding location could indicate functionality of miRNA:target
interaction or be effective on abundance of targets(TOFIX abundance
etkisi derken, bazen azaltabilir, bazen arttırabilir, manasında mı?,
evet hocam anlaşılmıyorsa değiştirebilirim). It has been shown that
miRNAs exhibit repressive activity via, 6-8 nt long sequence that is
perfectly complementary with targets, seed region at the 5' end of
miRNAs (Bartel 2009; Grimson et al. 2007). On the other hand, some
researchers have reported that seed sequence of miRNA can have
mismatches or bulged/wobble nucleotides and may locate in region other
than 5'end of miRNAs (Hafner et al. 2010; Helwak et al. 2013). On top of
all these factors, abundance of miRNAs and targets and miRNA:target
ratio in cells predominantly affect efficiency of miRNA:target
interaction (Arvey et al. 2010; Bosson, Zamudio, and Sharp 2014; Denzler
et al. 2014).

As it is possible for miRNAs to suppress multiple targets, an individual
mRNA molecule can also be targeted by multiple miRNAs. In that case, the
targeted mRNAs exhibit competitor behavior, that is hypothesized as
competing endogenous RNAs (ceRNAs) (Ala et al. 2013; Cesana and Daley
2013), against their miRNAs. Briefly, Ala et al.~have explained the
ceRNA hypothesis as disturbance of the other target when one of the
targets on a steady-state system that included one miRNA and two target
was perturbed with expression change (Ala et al. 2013). Regarding
interaction between miRNAs and their target in a cell, explaining and
predicting results of an individual perturbation is difficult due to
complexity of interactions. Various computational and experimental
studies have tackled the problem of unraveling ceRNA:miRNA interactions.
It has been observed that when abundance of one of the targets of
miR-122 was increased, the other target expression also slightly
increased as a result of decreasing repressive activity of miR-122 on
them (Denzler et al. 2014). Bosson et al.~have developed a mathematical
model for changes on total target pool concentration after grouping
targets according to affinity and demonstrated that miRNA activity
correlated with affinity between miRNA and target (Bosson, Zamudio, and
Sharp 2014). Cooperative efficiency of miRNAs as well as competitor
behaviors of targets were also studied and it has been demonstrated to
be crucial for regulating available mRNA levels of targets (Denzler et
al. 2016). MiRNA:target interactions have been modeled as stoichiometric
and catalytic mechanisms and Figliuzzi et al.~have recommended handling
models in network context (Figliuzzi, Marinari, and De Martino 2013).
The model that can explain miRNA target interaction through topological
features has been applied at bipartite network by Nitzan et al.~(Nitzan
et al. 2014). Robinson and Henderson applied the model that handles
miRNA:target direct and indirect interactions via common miRNA of gene
and target of miRNA, at bipartite network. It has been demonstrated that
all miRNAs and targets in the network can interact with each other
through common miRNAs and genes, without interaction between the same
type of nodes (Robinson and Henderson 2018).Associated genes that are
targets of the same miRNAs have been found with help of correlation of
gene expression changes in recent algorithm (Markus List 2017). List et
al.~have specified that their approach can be useful for ceRNA studies
and published their approach as an R package.

@@

\begin{figure}
\hypertarget{fig1}{%
\centering
\includegraphics{Fig1.jpg}
\caption{Schematic presentation of mechanism of network based model. a)
In steady state, miRNAs (M: gray triangle) repress targets (G: circle
coconut) according to proportion of their targets' expression. b) Two
fold increase in transcript level of Gene2 (G2) acts as a trigger. c)
Distribution of miRNA1 (M1) changes. d) The change at expression of
common target effects changes of proportional distribution of miRNA2
(M2). Expression values are rounded to integers for
simplicity}\label{fig1}
}
\end{figure}

@@

The repression efficiency of a miRNA on the individual target
(\(Eff_{gi}\)) is calculated according to equation \eqref{eq:1}; where
miRNA expression (\(C_m\)) in local neighborhood is distributed among
targets using individual gene expression levels (\(C_{gi}\)) (\#TODO
isn't \(C_g\) gene expressions in the group is equal to total of
\(C_{gi}\) in group?). For the genes targeted by multiple miRNAs,
cooperative activity of miRNAs, \(R\), is calculated by summing
repression activity of each miRNA (Equation \eqref{eq:2}).

\begin{equation} 
    Eff_{gi}= C_m \times C_{gi}/\sum_{1}^{i} C_g \tag{1}\label{eq:1}
\end{equation} \begin{equation}
   R_{gi}= Eff_{i1} + Eff_{i2} ... \tag{2}\label{eq:2}
\end{equation}

@@

Interactions between miRNAs and their targets can be affected from
various factors. So, our model integrates multiple factors when
calculating overall miRNA activity. We classified factors into two
categories. Factors effecting binding determine interaction between
miRNA and target and they alter amount of miRNA sequestered to target.
Factors effecting efficiency determine proportion of target mRNA
degraded by bound miRNA (\#TODO fix this sentence). In other words,
binding factors exert their influence before or during binding,
efficiency factors exert their influence after binding. In the
literature, binding free energy (Cao and Chen 2012 ; Helwak et al. 2013)
and seed type (Werfel et al. 2017) in miRNA:target interactions are
described as factors effecting binding affinity. Efficiency factors
determine how many of miRNA:target complexes will result in inhibition
and binding region on target drastically effect miRNA degradation
efficiency (Hausser et al. 2013; Helwak et al. 2013). Both binding and
efficiency factors are normalized to their maximum values and scaled to
{[}0,1{]} interval. The normalized values of factors take into account
to determine binding activity and miRNA efficiency on targets
(\autoref{fig2}). Binding affinities (activity, \(Eff\)) of miRNAs on
each individual gene are calculated as shown in equation \eqref{eq:4};
\(C_m\), miRNA expression in the group; \(C_g\), Gene expression;
\(C_{gi}\), individual gene expression; gi, individual gene; g, whole
genes in a group (\autoref{fig2}c). \begin{equation}
Eff_{gi}= C_m \times E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}/(\sum_{1}^{i} E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}) \tag{4}\label{eq:4}
\end{equation} \begin{equation}
Eff_{gi}= Eff_{gi}*RE^\prime_{gi} \tag{5}\label{eq:5}
\end{equation}

@@

After miRNA binds to its target, but might not repress to bound target.
The functionality of bound miRNA on target depends on efficiency factors
like region that is binding sequence of miRNA on its target. Exact
repression efficiency of miRNA is calculated according to equation
\eqref{eq:5} (\autoref{fig2}d); \(RE^\prime_{gi}\), normalized values of
region efficiency coefficient between miRNA and gene. The cooperative
repression activity of miRNAs to their common targets is figured out as
shown in \autoref{fig2}e.

@@

We have applied our model in a real dataset for which experimental
measurements of various factors were available. Expression levels of
miRNA and genes in tumor and normal tissue of single patient are
retrieved from TCGA(\#TODO cite). High-throughput experimental datasets
which are provided miRNA:gene target pairs with interaction factors
(Helwak et al. 2013; Moore et al. 2015). We have combined miRNA and gene
expression datasets via miRNA:target gene dataset retrieved from
\ldots{} (\#TODO name of database and its
citation)-\textgreater{}suplementary datasets of (Helwak et al. 2013;
Moore et al. 2015) . Detailed description of network construction and
its code is available in Supplementary data (\#TODO burada link nasıl
olacak, word dosyası veya PDF dosyası ismi mi vermemiz mi gerekiyor,
yoksa sadece Sup Data denilmesi yeterli mi? ilgili dosya:
TCGA\_E9-A1N5\_article.Rmd). ABCC1 gene, over-expressed in tumor tissue,
was selected as trigger for simulation of integrated dataset. After
simulation of network, we have compared simulation results and tumor
tissue expression levels.

@@

We have developed a network-based approach to assess effects of
expression level changes in competitive ceRNA regulation. The basic mode
of miRNA repression activity has been based on miRNA and target
abundance in various researches(Arvey et al. 2010; Denzler et al. 2014).
Our approach can effortlessly calculate effects of expression changes
when abundance levels of miRNAs and targets is only available factor. In
sample network given in Figure (\autoref{fig1}), after an increase in
expression level of a gene (G2), expression values of other genes also
changed due to redistribution of miRNA among its targets. Previous
studies have shown that if a gene abundance increases in ceRNA system,
expression levels of genes targeted by shared miRNA are also affected
(Lai, Wolkenhauer, and Vera 2016; Salmena et al. 2011; Tay, Rinn, and
Pandolfi 2014). It was observed that primary neighborhoods of the
trigger gene change with the same ratio, but cooperative efficiency of
miRNAs on common target causes differently change in common target
(\#TODO last sentence needs some clarification, after ``but
cooperative\ldots{}'') (\autoref{fig1}d). Genes targeted by multiple
miRNAs act as a trigger for adjacent local neighborhood of targeting
miRNAs, causing changes in expression levels of genes outside the local
neighborhood of original trigger gene. Therefore, primary expression
level change in gene (G2) causes changes in other group of genes (G5 and
G6) even though original trigger gene (G2) and genes in other group are
not targeted by common miRNA. In addition, as shown by ceRNA hypothesis
model of Ala et al., after the increase of gene expression level of G2,
the miRNA that is found in the same group (M1) tended to be less
repressive on its remaining targets (G1, G3 and G4). It's important to
note that the changes in gene expression levels will have more
pronounced effect if miRNA:target ratio is high, i.e., more miRNA
available per target, which was reported in previous findings (Arvey et
al. 2010; Bosson, Zamudio, and Sharp 2014; Denzler et al. 2014).
\href{}{}

@@

\href{}{\#TODO}: \textless{}\textgreater{} ``title should be more
informative''-\textgreater{}

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Network based multifactorial modelling microRNA-target interaction.},
            pdfauthor={Selcen Ari; Alper Yilmaz},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Network based multifactorial modelling microRNA-target interaction.}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Selcen Ari\footnote{Department of Bioengineering, Yildiz Technical
  University, Istanbul, Turkey} \\ Alper Yilmaz\footnote{Department of Bioengineering, Yildiz Technical
  University, Istanbul, Turkey}}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{21 Ağustos 2019}

@@

Competing endogenous RNA (ceRNA) regulations and crosstalk between
various types of non-coding RNA in human is remarkable in means of miRNA
regulation. Many studies have pointed out that an alteration in
miRNA:target interaction can result in unexpected changes due to
indirect and complex interactions. In this paper, we defined a new
network-based model that handles miRNA:ceRNA interactions with
expression values. Our model is able to handle miRNA interaction factors
such as seed type, binding energy, if provided. Our approach is able to
reveal that a perturbation in an element of network affects whole
competing elements differently and cooperative efficiencies of miRNAs on
common targets could be calculated. Our findings emphasized importance
of miRNA:target ratios being crucial, as reported by previous studies.
We have showed that the competing elements which have the same or close
expression values may not be affected equally from the perturbation
because of repression functionality depended on interaction factors of
miRNA target pairs (answer: Hocam etkileşim faktörlerinden dolayı
ekspresyonları yakın olsa bile baskılanma aktitesinin aynı
sergilenmeyeceğini ifade etmeye çalıştım. Bu şekilde anlaşılıyor mu?).
We applied the model to real sample consisting of breast cancer gene and
miRNA expression dataset and experimental miRNA:target interaction
dataset all generated via high throughput sequencing methods. A gene
over-expressed in tumor tissue, namely \emph{ABCC1}, is used as
perturbing element. We have observed that change in expression level of
single gene in miRNA:target network is sufficient to perturb regulations
in whole network, due to unforeseen and unpredicted regulation which are
only visible when considered in network context. Therefore, this model
helps unveiling the crosstalk between elements in miRNA:target network
where abundance of target and sponge effect are taken into account. The
model is scalable and can be plugged in with emerging miRNA effectors
such as circRNAs. The model is available as R package
\href{https://github.com/selcenari/ceRNAnetsim}{ceRNAnetsim}.

@@

MicroRNAs (miRNAs) are a family of short non-coding RNAs which are key
regulator of gene expression through various post-transcriptional
mechanisms. Although the mechanisms by which miRNA represses are not
fully understood, miRNAs predominantly repress their targets. Repressive
activities of miRNAs vary depending on many factors that are significant
to microRNA:target interactions. These factors include microRNA:target
binding energy, binding location in target sequence, base pairing types
between miRNA and target, abundance of miRNAs and targets (Grimson et
al. \protect\hyperlink{ref-grimson_microrna_2007}{2007}). Binding
energies of miRNA:target complexes vary based on nucleotide context and
determine folding stability of complex (Cao and Chen
\protect\hyperlink{ref-cao_predicting_2012}{2012}). It has been
demonstrated that the binding energy between miRNA and target indicates
stability or affinity of complex (Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}) and does not directly
determine repressive activity of miRNA (Cao and Chen
\protect\hyperlink{ref-cao_predicting_2012}{2012}). Early studies have
argued that 2-8 nt sequence, seed, located in miRNA 5'end bind to
specific sequence located in 3'UTR of its target (Bartel
\protect\hyperlink{ref-bartel_micrornas_2004}{2004}; Lewis, Burge, and
Bartel \protect\hyperlink{ref-lewis_conserved_2005}{2005}). In recent
studies, it has been shown that miRNAs can interact with targets via
sequences located in regions such as 5'UTR or CDS (Hausser et al.
\protect\hyperlink{ref-hausser_analysis_2013}{2013}; Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}). These studies
also showed that binding location could indicate functionality of
miRNA:target interaction or be effective on abundance of targets(TOFIX
abundance etkisi derken, bazen azaltabilir, bazen arttırabilir,
manasında mı? answer: evet hocam anlaşılmıyorsa değiştirebilirim). It
has been shown that miRNAs exhibit repressive activity via, 6-8 nt long
sequence that is perfectly complementary with targets, seed region at
the 5' end of miRNAs (Bartel
\protect\hyperlink{ref-bartel_micrornas:_2009}{2009}; Grimson et al.
\protect\hyperlink{ref-grimson_microrna_2007}{2007}). On the other hand,
some researchers have reported that seed sequence of miRNA can have
mismatches or bulged/wobble nucleotides and may locate in region other
than 5'end of miRNAs (Hafner et al.
\protect\hyperlink{ref-hafner_transcriptome-wide_2010}{2010}; Helwak et
al. \protect\hyperlink{ref-helwak_mapping_2013}{2013}). On top of all
these factors, abundance of miRNAs and targets and miRNA:target ratio in
cells predominantly affect efficiency of miRNA:target interaction (Arvey
et al. \protect\hyperlink{ref-arvey_target_2010}{2010}; Bosson, Zamudio,
and Sharp \protect\hyperlink{ref-bosson_endogenous_2014}{2014}; Denzler
et al. \protect\hyperlink{ref-denzler_assessing_2014}{2014}).

As it is possible for miRNAs to suppress multiple targets, an individual
mRNA molecule can also be targeted by multiple miRNAs. In that case, the
targeted mRNAs exhibit competitor behavior, that is hypothesized as
competing endogenous RNAs (ceRNAs) (Ala et al.
\protect\hyperlink{ref-ala_integrated_2013}{2013}; Cesana and Daley
\protect\hyperlink{ref-cesana_deciphering_2013}{2013}), against their
miRNAs. Briefly, Ala et al.~have explained the ceRNA hypothesis as
disturbance of the other target when one of the targets on a
steady-state system that included one miRNA and two target was perturbed
with expression change (Ala et al.
\protect\hyperlink{ref-ala_integrated_2013}{2013}). Regarding
interaction between miRNAs and their target in a cell, explaining and
predicting results of an individual perturbation is difficult due to
complexity of interactions. Various computational and experimental
studies have tackled the problem of unraveling ceRNA:miRNA interactions.
It has been observed that when abundance of one of the targets of
miR-122 was increased, the other target expression also slightly
increased as a result of decreasing repressive activity of miR-122 on
them (Denzler et al.
\protect\hyperlink{ref-denzler_assessing_2014}{2014}). Bosson et
al.~have developed a mathematical model for changes on total target pool
concentration after grouping targets according to affinity and
demonstrated that miRNA activity correlated with affinity between miRNA
and target (Bosson, Zamudio, and Sharp
\protect\hyperlink{ref-bosson_endogenous_2014}{2014}). Cooperative
efficiency of miRNAs as well as competitor behaviors of targets were
also studied and it has been demonstrated to be crucial for regulating
available mRNA levels of targets (Denzler et al.
\protect\hyperlink{ref-denzler_impact_2016}{2016}). MiRNA:target
interactions have been modeled as stoichiometric and catalytic
mechanisms and Figliuzzi et al.~have recommended handling models in
network context (Figliuzzi, Marinari, and De Martino
\protect\hyperlink{ref-figliuzzi_micrornas_2013}{2013}). The model that
can explain miRNA target interaction through topological features has
been applied at bipartite network by Nitzan et al.~(Nitzan et al.
\protect\hyperlink{ref-nitzan_interactions_2014}{2014}). Robinson and
Henderson applied the model that handles miRNA:target direct and
indirect interactions via common miRNA of gene and target of miRNA, at
bipartite network. It has been demonstrated that all miRNAs and targets
in the network can interact with each other through common miRNAs and
genes, without interaction between the same type of nodes (Robinson and
Henderson
\protect\hyperlink{ref-robinson_modelling_2018}{2018}).Associated genes
that are targets of the same miRNAs have been found with help of
correlation of gene expression changes in recent algorithm (Markus List
\protect\hyperlink{ref-markus_list_sponge_2017}{2017}). List et al.~have
specified that their approach can be useful for ceRNA studies and
published their approach as an R package.

@@

\begin{figure}
\hypertarget{fig1}{%
\centering
\includegraphics{Fig1.jpg}
\caption{Schematic presentation of mechanism of network based model. a)
In steady state, miRNAs (M: gray triangle) repress targets (G: circle
coconut) according to proportion of their targets' expression. b) Two
fold increase in transcript level of Gene2 (G2) acts as a trigger (shown
with red in figure). c) Distribution of miRNA1 (M1) changes. d) The
change at expression (shown with light red because minor changes are
occured on their expression values.) of common target effects changes of
proportional distribution of miRNA2 (M2). Expression values are rounded
to integers for simplicity}\label{fig1}
}
\end{figure}

@@

The repression efficiency of a miRNA on the individual target
(\(Eff_{gi}\)) is calculated according to equation \eqref{eq:1}; where
miRNA expression (\(C_m\)) in local neighborhood is distributed among
targets using individual gene expression levels (\(C_{gi}\)) (\#FIX
isn't \(C_g\) gene expressions in the group is equal to total of
\(C_{gi}\) in group? answer: its my false. now its corect.). For the
genes targeted by multiple miRNAs, cooperative activity of miRNAs on a
target gene, \(R\), is calculated by summing repression activity of each
miRNA (Equation \eqref{eq:2}).

\begin{equation} 
    Eff_{gi}= C_m \times C_{gi}/\sum_{1}^{i} C_{gi} \tag{1}\label{eq:1}
\end{equation}

\begin{equation}
   R_{gi}= Eff_{i1} + Eff_{i2} ... \tag{2}\label{eq:2}
\end{equation}

@@

Interactions between miRNAs and their targets can be affected from
various factors. So, our model integrates multiple factors when
calculating overall miRNA activity. We classified factors into two
categories. Factors effecting binding determine interaction between
miRNA and target and they alter amount of miRNA sequestered to target.
Degradation efficiency factors assign amount of degradated target amount
in miRNA:target pairs.(\#FIX fix this sentence: answer:). In other
words, binding factors exert their influence before or during binding,
efficiency factors exert their influence after binding. In the
literature, binding free energy (Cao and Chen
\protect\hyperlink{ref-cao_predicting_2012}{2012} ; Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}) and seed type (Werfel
et al. \protect\hyperlink{ref-werfel_preferential_2017}{2017}) in
miRNA:target interactions are described as factors effecting binding
affinity. Efficiency factors determine how many of miRNA:target
complexes will result in inhibition and binding region on target
drastically effect miRNA degradation efficiency (Hausser et al.
\protect\hyperlink{ref-hausser_analysis_2013}{2013}; Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}). Both binding and
efficiency factors are normalized to their maximum values and scaled to
{[}0,1{]} interval. The normalized values of factors take into account
to determine binding activity and miRNA efficiency on targets
(\autoref{fig2}). Binding affinities (activity, \(Eff\)) of miRNAs on
each individual gene are calculated as shown in equation \eqref{eq:3};
\(C_m\), miRNA expression in the group; \(C_{gi}\), individual gene
expression; gi, individual gene (\autoref{fig2}c). \begin{equation}
Eff_{gi}= C_m \times E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}/(\sum_{1}^{i} E^\prime_{gi} \times STE^\prime_{gi} \times C_{gi}) \tag{3}\label{eq:3}
\end{equation}

\begin{equation}
Eff_{gi}= Eff_{gi}\times RE^\prime_{gi} \tag{4}\label{eq:4}
\end{equation}

@@

I shown in figure and explained at following. I didn't fully understand
whether it is enough or not.

After miRNA binds to its target, but might not repress to bound target.
The functionality of bound miRNA on target depends on efficiency factors
like region that is binding sequence of miRNA on its target. Exact
repression efficiency of miRNA is calculated according to equation
\eqref{eq:4} (\autoref{fig2}d); \(RE^\prime_{gi}\), normalized values of
region efficiency coefficient between miRNA and gene. The cooperative
repression activity of miRNAs to their common targets is figured out as
shown in \autoref{fig2}e.

@@

We have applied our model in a real dataset for which experimental
measurements of various factors were available. Expression levels of
miRNA and genes in tumor and normal tissue of single patient are
retrieved from
\href{https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga}{TCGA}(\#TODO).
High-throughput experimental datasets which are provided miRNA:gene
target pairs with interaction factors (Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}). We have combined
miRNA and gene expression datasets via miRNA:target gene dataset
retrieved from \ldots{} (\#FIX name of database and its citation)
answer-\textgreater{}suplementary datasets of (Helwak et al.
\protect\hyperlink{ref-helwak_mapping_2013}{2013}; Moore et al.
\protect\hyperlink{ref-moore_mirnatarget_2015}{2015}). Detailed
description of network construction and its code is available in
Supplementary data (\#TODO burada link nasıl olacak, word dosyası veya
PDF dosyası ismi mi vermemiz mi gerekiyor, yoksa sadece Sup Data
denilmesi yeterli mi? ilgili dosya: TCGA\_E9-A1N5\_article.Rmd). ABCC1
gene, over-expressed in tumor tissue, was selected as trigger for
simulation of integrated dataset. After simulation of network, we have
compared simulation results and tumor tissue expression levels.

@@

We have developed a network-based approach to assess effects of
expression level changes in competitive ceRNA regulation. The basic mode
of miRNA repression activity has been based on miRNA and target
abundance in various researches(Arvey et al.
\protect\hyperlink{ref-arvey_target_2010}{2010}; Denzler et al.
\protect\hyperlink{ref-denzler_assessing_2014}{2014}). Our approach can
effortlessly calculate effects of expression changes when abundance
levels of miRNAs and targets is only available factor. In sample network
given in Figure (\autoref{fig1}), after an increase in expression level
of a gene (G2), expression values of other genes also changed due to
redistribution of miRNA among its targets. Previous studies have shown
that if a gene abundance increases in ceRNA system, expression levels of
genes targeted by shared miRNA are also affected (Lai, Wolkenhauer, and
Vera \protect\hyperlink{ref-lai_understanding_2016}{2016}; Salmena et
al. \protect\hyperlink{ref-salmena_cerna_2011}{2011}; Tay, Rinn, and
Pandolfi \protect\hyperlink{ref-tay_multilayered_2014}{2014}). It was
observed that expression levels of primary neighborhoods which do not
interact with another miRNA of the trigger gene change in relation to
their expression. However, expression level change of common target (G4)
in network has been observed different from G1 that has the same
expression value with G4 target at initial conditions because more than
one miRNA repress this target (\autoref{fig1}d) (\#FIX last sentence
needs some clarification, after ``but cooperative\ldots{}'').

Genes targeted by multiple miRNAs act as a trigger for adjacent local
neighborhood of targeting miRNAs, causing changes in expression levels
of genes outside the local neighborhood of original trigger gene.
Therefore, primary expression level change in gene (G2) causes changes
in other group of genes (G5 and G6) even though original trigger gene
(G2) and genes in other group are not targeted by common miRNA. In
addition, as shown by ceRNA hypothesis model of Ala et al., after the
increase of gene expression level of G2, the miRNA that is found in the
same group (M1) tended to be less repressive on its remaining targets
(G1, G3 and G4). It's important to note that the changes in gene
expression levels will have more pronounced effect if miRNA:target ratio
is high, i.e., more miRNA available per target, which was reported in
previous findings (Arvey et al.
\protect\hyperlink{ref-arvey_target_2010}{2010}; Bosson, Zamudio, and
Sharp \protect\hyperlink{ref-bosson_endogenous_2014}{2014}; Denzler et
al. \protect\hyperlink{ref-denzler_assessing_2014}{2014}). \href{}{}

@@

\href{}{\#FIX}: \textless{}\textgreater{} ``title should be more
informative''-\textgreater{} answer.

Earlier studies reported that miRNA regulatory interactions are affected
by different parameters. For example, Xu et al.~have investigated the
importance of seed pairing type between miRNAs and their targets and
target site location by using proteomics dataset (Xu, Wang, and Liu
\protect\hyperlink{ref-xu_characterization_2014}{2014}). They have
proposed that the features of binding between miRNA and target can be
critical for miRNA efficiency. In addition, binding energy between miRNA
and targets is a significant determinant for miRNA efficiency and it has
been reported that strength of miRNA:target interactions is depended on
binding energy of complexes (Breda et al.
\protect\hyperlink{ref-breda_quantifying_2015}{2015}). Similarly,
another study revealed that affinity is correlated with seed pairing of
miRNA:target pairs and suggested that affinity is correlated with number
of canonical seed base pairing (Bosson, Zamudio, and Sharp
\protect\hyperlink{ref-bosson_endogenous_2014}{2014}). Therefore, we
have integrated aforementioned interaction parameters that could be
useful for predicting miRNA repression activity more accurately. The
sample dataset used in \autoref{fig1} is recalculated with additional
factors \autoref{tab:one} in effect with same trigger, two fold increase
in Gene2 (\autoref{fig3}). When the factors were taken into account in
the system, miRNA efficiencies varied as shown in \autoref{fig3}a.
Although the miRNA:target expression ratios in steady-state were same in
comparison with the sample dataset without factors, efficiency of
binding and repression have changed.
